Ludwig User Tutorial

### 2017-06-28 docs/tutorial/tutorial.tex
Edited tutorial section for test1
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3140 40876441-3496-0410-a613-a2fca35b2ae2
Added tutorial section for test2
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3144 40876441-3496-0410-a613-a2fca35b2ae2

\usepackage{amsmath}
\usepackage{moreverb}
\usepackage{lscape}
\usepackage{epic}
\usepackage[pdftex]{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{enumerate}
\graphicspath{{./tutorialFigs/}}

@@

\begin{enumerate}
\item Copy the input file \texttt{input} located in the directory \texttt{/trunk/docs/tutorial/test1} 
to your working directory. Note that this input file is a modified version of the 
reference input file \texttt{input.ref} located in \texttt{/trunk/src} with a limited set of parameters
for this run. 
\item Run the code in your working directory by issuing the command:
\begin{lstlisting}
$ mpirun -np 4 ./Ludwig.exe input
\end{lstlisting}
\item You should now see statistics being reported on your standard output and 
output files being created with names of the form \texttt{vel-XXXXXXXX.001-001}. 
These files contain the components of the velocity vectors in Cartesian coordinates at each lattice point in the simulation. 
\item The first step to the post-processing of these data is to copy the files \texttt{vtk\_extract.c} 
and \texttt{vtk\_extract\_script.py} from \texttt{/trunk/util} into your working directory.
\item Edit \texttt{vtk\_extract.c}. Specifically, set the flags for:
\begin{enumerate}
\item \texttt{input\_binary\_=0} \\ if the velocity data is in ASCII format (see flag in input file),
\item \texttt{output\_binary\_=0} \\ as the output of the postprocessing routines should also be in ASCII format,
\item \texttt{output\_index=1} \\ to obtain Cartesian lattice indices \textit{i, j, k}, 
\item \texttt{vtk\_header=0} \\ to omit headers (these will be required later for visualisation in \textit{ParaView})
\item \texttt{output\_cmf=0} to produce data in row-major format (column-major format is required for \textit{ParaView}).
\end{enumerate}
\item Edit \texttt{vtk\_extract\_script.py}. Set:
\begin{enumerate}
\item \texttt{nstart=1000}, \texttt{nint=1000} and \texttt{nend=10000} to set the range and increments of the velocity data
to be processed.
\item \texttt{vel=1} for velocity post-processing, 
\item The other flags should be set to zero, e.g. \texttt{phi=0}. 
\end{enumerate}
\item You should now see files named \texttt{vel-XXXXXXXX.vtk} (Note: The extension \texttt{.vtk} is only used here for simplicity.
It does not mean the generated data is in vtk-format). 
The first three columns of these ASCII files are the spatial lattice indices followed by the Cartesian components of the velocity vectors.\\

To plot the velocity profile of the system, you can plot the $y$-component of the velocity against 
the spatial $x$-coordinate on the lattice with the command:
\begin{lstlisting}
gnuplot> p`vel-XXXXXXXX.vtk' u 1:5
\end{lstlisting} 
By plotting the data in each file, you should be able to see the velocity profile converging to the final profile 
as shown in Fig. \ref{fig:velocityProfile}. In the directory \texttt{/trunk/docs/tutorial/test1} a reference file of the final 
output at time step $10,000$ can be found, in addition to a file containing the standard output of the code.
\end{enumerate}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{velProf.png}
  \caption{The parabolic velocity profile of a simple fluid in a cavity after $10,000$ time-steps}.
  \label{fig:velocityProfile}
  \end{center}
\end{figure}

>>>>

\usepackage{amsmath}
\usepackage{moreverb}
\usepackage{lscape}
\usepackage{epic}
\usepackage[pdftex]{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{float}
\graphicspath{{./tutorialFigs/}}

@@

\begin{enumerate}
\item Copy the input file \texttt{input} located in the directory \texttt{/trunk/docs/tutorial/test1} 
to your working directory. Note that this input file is a modified version of the 
reference input file \texttt{input.ref} located in \texttt{/trunk/src} with a limited set of parameters
for this run. 
\item Run the code in your working directory by issuing the command:
\begin{lstlisting}
$ mpirun -np 4 ./Ludwig.exe input
\end{lstlisting}
\item You should now see statistics being reported on your standard output and 
output files being created with names of the form \texttt{vel-XXXXXXXX.001-001}. 
These files contain the components of the velocity vectors in Cartesian coordinates at each lattice point in the simulation. 
\item The first step to the post-processing of these data is to copy the files \texttt{vtk\_extract.c} 
and \texttt{vtk\_extract\_script.py} from \texttt{/trunk/util} into your working directory.
\item Edit \texttt{vtk\_extract.c}. Specifically, set the flags for:
\begin{enumerate}
\item \texttt{input\_binary\_=0} \\ if the velocity data is in ASCII format (see flag in input file),
\item \texttt{output\_binary\_=0} \\ as the output of the postprocessing routines should also be in ASCII format,
\item \texttt{output\_index=1} \\ to obtain Cartesian lattice indices \textit{i, j, k}, 
\item \texttt{vtk\_header=0} \\ to omit headers (these will be required later for visualisation in \textit{ParaView}),
\item \texttt{output\_cmf=0} to produce data in row-major format (column-major format is required for \textit{ParaView}).
\end{enumerate}
\item Edit \texttt{vtk\_extract\_script.py}. Set:
\begin{enumerate}
\item \texttt{nstart=1000}, \texttt{nint=1000} and \texttt{nend=10000} to set the range and increments of the velocity data
to be processed.
\item \texttt{vel=1} for velocity post-processing, 
\item The other flags should be set to zero, e.g. \texttt{phi=0}. 
\end{enumerate}
\item Issue the command:
\begin{lstlisting}
$ python vtk_extract_script.py
\end{lstlisting}
\item You should now see files named \texttt{vel-XXXXXXXX.vtk} (Note: The extension \texttt{.vtk} is only used here for simplicity.
It does not mean the generated data is in vtk-format). 
The first three columns of these ASCII files are the spatial lattice indices followed by the Cartesian components of the velocity vectors.\\

To plot the velocity profile of the system, you can plot the $y$-component of the velocity against 
the spatial $x$-coordinate on the lattice with the command:
\begin{lstlisting}
gnuplot> p`vel-XXXXXXXX.vtk' u 1:5
\end{lstlisting} 
By plotting the data in each file, you should be able to see the velocity profile converging to the final profile 
as shown in Fig. \ref{fig:velocityProfile}.

In the directory \texttt{/trunk/docs/tutorial/test1} a reference file of the final 
output at time step $10,000$ can be found, in addition to a file containing the standard output of the code.
\end{enumerate}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{velProf.png}
  \caption{The parabolic velocity profile of a simple fluid in a cavity after $10,000$ time-steps.}
  \label{fig:velocityProfile}
  \end{center}
\end{figure}

\subsection{Test 1: Spinodal Decomposition of a Binary Fluid}

This section outlines the second example, the spinodal decomposition of a binary fluid in 
three-dimensions.
The example describes the post-processing procedure for this system and a brief guide to 
the visualisation of a 3D system using ParaView.
As in the previous example, the code should be compiled in parallel and the executable 
\texttt{Ludwig.exe} should be copied into your working directory.

\begin{enumerate}
\item Copy the input file \texttt{input} located in the directory 
\texttt{/trunk/docs/tutorial/test2} 
to your working directory. This input file contains the parameters required for this run. 
\item Run the code in your working directory by issuing the command:
\begin{lstlisting}
$ mpirun -np 8 ./Ludwig.exe input
\end{lstlisting}
\item As before, you should now see statistics being reported on your standard output but 
in this case two sets of output files being created containing the velocity and compositional
order parameter ($\phi$) data at each lattice point in the simulation. In this example, these 
data are in binary format.
\item The post-processing of these data again requires the files \texttt{vtk\_extract.c} 
and \texttt{vtk\_extract\_script.py} from \texttt{/trunk/util}.
\item Edit \texttt{vtk\_extract.c}. Specifically, set the flags for:
\begin{enumerate}
\item \texttt{input\_binary\_=1} \\ if the velocity and $\phi$ data are in binary format (see 
flag in input file),
\item \texttt{output\_binary\_=0} \\ as the output of the post-processing routines should be 
in ASCII format,
\item \texttt{output\_index=0} \\ as Cartesian lattice indices are not required, 
\item \texttt{vtk\_header=1} \\ to include headers,
\item \texttt{output\_cmf=1} to produce data in column-major format.
\end{enumerate}
\item Edit \texttt{vtk\_extract\_script.py}. Set:
\begin{enumerate}
\item \texttt{nstart=1000}, \texttt{nint=1000} and \texttt{nend=10000} to set the range and 
increments of the velocity data
to be processed.
\item \texttt{vel=1} for velocity post-processing, 
\item \texttt{phi=1} for $\phi$ post-processing,
\item The other flags should be set to zero. 
\end{enumerate}
\item Issue the command:
\begin{lstlisting}
$ python vtk_extract_script.py
\end{lstlisting}
\item You should now see files with the extension \texttt{.vtk} corresponding to the velocity 
and $\phi$ data ready to be visualised using ParaView.
\begin{enumerate}
\item Open the files containing the $\phi$ data.
\item Click on the group of files in the pipeline browser and apply a contour map using 
Filters/Common/Contour or in the tool bar:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\linewidth]{contour.png}
  \caption{The contour button in the toolbar.}
  \label{fig:contour}
  \end{center}
\end{figure}

You should now see a three-dimensional rendering of the binary fluid.
\item Open the files containing the velocity data.
\item Click on the group of files in the pipeline browser and apply the glyph filter using 
Filters/Common/Glyph or in the tool bar:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\linewidth]{glyph.png}
  \caption{The glyph button in the toolbar}
  \label{fig:glyph}
  \end{center}
\end{figure}

\item Change the colour code to GlyphVector Magnitude.
\item In the pipeline browser, you should see:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.6\linewidth]{pipeline.png}
  \caption{The complete pipline browser for the visualisation of spinodal decomposition.}
  \label{fig:pipeline}
  \end{center}
\end{figure}

A 3D visualisation of the system should also now be shown:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.6\linewidth]{system.png}
  \caption{A visualisation of a separating binary fluid using ParaView.}
  \label{fig:sysVisualisation}
  \end{center}
\end{figure}

\item You can move through different frames of the simulation with the arrow buttons:

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.4\linewidth]{arrows.png}
  \caption{A visualisation of a separating binary fluid using ParaView.}
  \label{fig:sysVisualisation}
  \end{center}
\end{figure}

as you do this, you should see the coarsening of the binary fluid.

\item In the File menu you can save a screen shot of the visualisation or save the state of the system in order to reload the entire visualisation.
\end{enumerate}
\end{enumerate}
